### 2022-05-06 All_EDA.Rmd
Thesis Code Backup

---
output:
  html_document:
    theme: flatly
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 4
    df_print: paged
    code_folding: show
---

```{r, include=FALSE}
library(tidyverse)
library(tidytext)
library(scales)
library(AER)

options(scipen = 20)

knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.align = "center",
                      fig.width = 8)

```

```{r}
## Laptop Data

files <- dir()

files <- files[str_detect(dir(), "all_laptop_reviews.+")]

laptop_data <- map_dfr(files, read_csv)

laptop_data <- laptop_data %>% 
  filter(!Product_Name %in% c(
    'Samsung - Galaxy Book Pro 360 15.6" AMOLED Touch-Screen Laptop - Intel Evo Platform Core i7 - 16GB Memory - 1TB SSD - Mystic Navy',
    'ASUS - 14.0" Laptop - Intel Celeron N4020 - 4GB Memory - 64GB eMMC - Star Black - Star Black',
    'ASUS - 14" Chromebook - Intel Celeron N3350 - 4GB Memory - 32GB eMMC - Silver',
    'Samsung - Galaxy 13.3" 4K Ultra HD Touch-Screen Chromebook - Intel Core i5 - 8GB Memory - 256GB SSD - Mercury Gray'
  )) %>% 
  mutate(Product_Type = "Laptop")

laptop_data <- laptop_data %>%
  mutate(Total_Vote = Helpful + Unhelpful) %>% 
  filter(Total_Vote > 4) %>% 
  mutate(Helpfulness = Helpful / Total_Vote) %>% 
  mutate(is_Exp = 0)
```

```{r}
movies_data <- read_csv("homecoming15-02-22.csv") %>% 
  mutate(Movie = "Homecoming") %>% 
  bind_rows(read_csv("far_from_home15-02-22.csv") %>% 
              mutate(Movie = "Far From Home")) %>% 
  bind_rows(read_csv("no_way_home19-02-22.csv") %>% 
              mutate(Movie = "No Way Home"))
```

```{r}
movies_data <- movies_data %>% 
  filter(!is.na(Rating))

movies_data <- movies_data %>% 
  separate(Rating, into = c("Rating", "Total"), sep = "/", ) %>% 
  select(-Total) %>% 
  filter(Total_Vote > 4) %>% 
  mutate(Rating = as.double(Rating)) %>% 
  mutate(Helpfulness = Helpful_Vote / Total_Vote) %>% 
  mutate(Rating = Rating/2) %>% 
  mutate(is_Exp = 1)

# In this case I am removing every duplicated review.

movies_data <- movies_data %>% 
  mutate(Rev_dup = duplicated(Review) | duplicated(Review, fromLast = TRUE)) %>% 
  filter(Rev_dup == 0)
```

```{r}
laptop_data %>% names
```

```{r}
movies_data %>% names
```

```{r}
all_data <- laptop_data %>% 
  rename(Helpful_Vote = Helpful) %>% 
  select(-Unhelpful, -Product_Name, -Page) %>% 
  bind_rows(movies_data %>% select(-Movie))
```

```{r}
all_data <- all_data %>% 
  mutate(Doc_ID = row_number())
```

```{r}
all_data <- all_data %>% 
  mutate(Review = str_to_lower(Review, locale = "en"))
```

```{r}
words_agg <- all_data %>% 
  unnest_tokens(output = word, input = Review, to_lower = FALSE) %>% 
  anti_join(stop_words) %>% 
  count(Doc_ID)
```

```{r}
all_data <- words_agg %>% 
  inner_join(all_data, by = "Doc_ID") 
```

```{r}
summary(lm(Helpfulness ~ Rating*is_Exp + I(Rating^2)*is_Exp + log10(n)*is_Exp + log10(Total_Vote), data = all_data))
```

```{r}
summary(tobit(Helpfulness ~ Rating*is_Exp + I(Rating^2)*is_Exp + log10(n)*is_Exp + log10(Total_Vote), data = all_data, left = 0, right = 1))
```

```{r}
all_data %>% 
  ggplot(aes(n, Helpfulness))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~is_Exp)
```