### 2022-05-06 Movies_EDA.Rmd

---
output:
  html_document:
    theme: flatly
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 4
    df_print: paged
    code_folding: show
---

```{r, include=FALSE}
library(tidyverse)
library(tidytext)
library(scales)
library(AER)
library(ggfortify)
library(lubridate)
library(koRpus.lang.en)

options(scipen = 20)

knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.align = "center",
                      fig.width = 8)

```

```{r}
files <- dir()

files <- files[str_detect(files, "movie_.+")]

#movies_data <- movies_data %>% 
#  bind_rows(map_dfr(files, read_csv))

movies_data <- map_dfr(files, read_csv, show_col_types = FALSE)
```

```{r}
movies_data <- movies_data %>% 
  filter(!is.na(Rating))

movies_data <- movies_data %>% 
  #filter(Movie != "Midsommar") %>% 
  separate(Rating, into = c("Rating", "Total"), sep = "/", ) %>% 
  select(-Total) %>% 
  filter(Total_Vote > 4) %>% 
  mutate(Rating = as.double(Rating)) %>% 
  mutate(Helpfulness = Helpful_Vote / Total_Vote) %>% 
  mutate(Rating = Rating/2)
```

```{r}
## Removing All the Duplicated Reviews including the original
movies_data <- movies_data %>% 
  mutate(Rev_dup = duplicated(Review) | duplicated(Review, fromLast = TRUE)) %>% 
  filter(Rev_dup == 0)
```

```{r}
## Word Count

movies_data <- movies_data %>% 
  mutate(Doc_ID = row_number())
## Words
movies_agg <- movies_data %>% 
  mutate(Review = str_to_lower(Review, locale = "en")) %>% 
  unnest_tokens(output = word, input = Review, to_lower = FALSE) %>% 
  #anti_join(stop_words) %>% 
  count(Doc_ID, name = "Word_Count")

movies_data <- movies_data %>% 
  inner_join(movies_agg, by = "Doc_ID")

## Sentences

movies_sentence_agg <- movies_data %>% 
  unnest_tokens(output = Sentence, input = Review, to_lower = FALSE, token = "sentences") %>% 
  count(Doc_ID,name = "n_Sentence")

movies_data <- movies_data %>% 
  inner_join(movies_sentence_agg, by = "Doc_ID")

## Emotion Sentiments (EmoLex)

sentiments_agg <- movies_data %>%
  mutate(Review = str_to_lower(Review, locale = "en")) %>% 
  unnest_tokens(output = word, input = Review, to_lower = FALSE) %>% 
  inner_join(get_sentiments("nrc") %>% filter(word != "spider")) %>% 
  count(Doc_ID, sentiment, name = "Sent_Count") %>% 
  pivot_wider(names_from = sentiment, values_from = Sent_Count, values_fill = 0)

movies_data <- movies_data %>% 
  full_join(sentiments_agg, by = "Doc_ID")

variables <- names(sentiments_agg)[-1]

movies_data <- movies_data %>% 
  mutate(across(.cols = all_of(variables), .fns = ~ifelse(is.na(.), 0, .) ))

movies_data <- movies_data %>% 
  mutate(across(.cols = all_of(variables), .fns = function(.) ./Word_Count*100))

movies_data <- movies_data %>% 
  mutate(across(.cols = variables, .fns = function(.) log10(. + 1)))

```

```{r}
# 
# review_vec <-  movies_data %>%
#   pull(Review)
# 
# 
# SMOG_grade <- vector(mode = "double")
# SMOG_age <- vector(mode = "double")
# Flesch_raw <- vector(mode = "double")
# ARI_grade <- vector(mode = "double")
# 
# ## ARI Index
# # for (i in seq_len(length(review_vec))) {
# #   ARI_Index <- readability(tokenize(review_vec[i], lang = "en", format = "obj"),index = "ARI", quiet = TRUE)
# #
# #   ARI_grade <- append(ARI_grade, summary(ARI_Index) %>% pull(grade))
# # }
# 
# # length(ARI_grade)
# 
# ## Flesch Ease of Reading
# 
# # for (i in seq_len(length(review_vec))) {
# #   flesch_reading_ease <- readability(tokenize(review_vec[i], lang = "en", format = "obj"),index = "Flesch", quiet = TRUE)
# #
# #   Flesch_raw <- append(Flesch_raw, summary(flesch_reading_ease) %>% pull(raw))
# # }
# #
# # length(Flesch_raw)
# 
# ## SMOG
# 
# capture.output(
# for (i in seq_len(length(review_vec))) {
#   SMOG <-  readability(tokenize(review_vec[i], lang = "en", format = "obj"),index = "SMOG")
# 
#   SMOG_grade <- append(SMOG_grade, summary(SMOG) %>% pull(grade))
#   SMOG_age <-  append(SMOG_age, summary(SMOG) %>% pull(age))
# }
# ,file = nullfile()
# )
# length(SMOG_grade)
```

```{r}
# movies_data <- movies_data %>% 
#   bind_cols(tibble(
#     SMOG_grade = as.double(SMOG_grade),
#     SMOG_age = as.double(SMOG_age)))

# movies_data <- movies_data %>% 
#   bind_cols(tibble(
#     Flesch_raw = as.double(Flesch_raw)
#   ))
# 
# movies_data <- movies_data %>% 
#   bind_cols(tibble(
#     ARI_grade = as.double(ARI_grade)
#   ))
```

```{r}
## Timeline

movies_data <- movies_data %>% 
  mutate(Date = dmy(Date, locale = "english"))

first_dates <- movies_data %>% 
  group_by(Movie) %>% 
  slice_min(order_by = Date, n = 1, with_ties = F) %>% 
  select(Movie, Date) %>% 
  rename(First_Date = Date)

movies_data <- full_join(movies_data, first_dates, by = "Movie") %>% 
  mutate(Timeline = as.integer(Date - First_Date))
```

```{r}
movies_data %>% 
  ggplot(aes(Helpfulness))+
  geom_histogram(color = "white")

movies_data %>% 
  ggplot(aes(Rating))+
  geom_bar()

movies_data %>% 
  ggplot(aes(Word_Count))+
  geom_histogram(color = "white")+
  scale_x_log10()

movies_data %>% 
  ggplot(aes(n_Sentence))+
  geom_histogram(color = "white")+
  scale_x_log10()

movies_data %>% 
  ggplot(aes(Total_Vote))+
  geom_histogram(color = "white")+
  scale_x_log10()

```

```{r}
movies_data %>% 
  ggplot(aes(Rating, Helpfulness))+
  geom_point()+
  geom_smooth()

movies_data %>% 
  ggplot(aes(Rating, Helpfulness))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~Movie)

movies_data %>% 
  ggplot(aes(Timeline, Helpfulness))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~Movie, scales = "free_x")

movies_data %>% 
  ggplot(aes(log10(Word_Count), Helpfulness))+
  geom_point(alpha = 0.3)+
  geom_smooth()

movies_data %>% 
  ggplot(aes((n_Sentence), Helpfulness))+
  geom_point(alpha = 0.3)+
  geom_smooth()+
  scale_x_log10()

movies_data %>% 
  ggplot(aes(log10(Word_Count), Helpfulness))+
  geom_point(alpha = 0.3)+
  geom_smooth()+
  facet_wrap(~Movie)

```

```{r}
movies_data %>% 
  ggplot(aes(Word_Count, Helpfulness))+
  geom_point(alpha = 0.2)+
  geom_smooth()+
  facet_wrap(~Rating)+
  scale_x_log10()

movies_data %>% 
  mutate(WC_Binned = cut_number(Word_Count, n = 10)) %>% 
  ggplot(aes(Rating, Helpfulness))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~WC_Binned)

movies_data %>% 
  mutate(across(.cols = variables, .fns = function(.) log10(. + 1))) %>% 
  select(variables, Word_Count,Rating, Total_Vote, Timeline) %>%
  GGally::ggcorr(label = T)
```

```{r}
mod_lm <- lm(Helpfulness ~ Rating + I(Rating^2) + I(Rating^3)+ log10(Word_Count) + log10(Total_Vote)  + log10(Word_Count)*Rating + Timeline + anticipation   + positive + negative + trust,
           data = movies_data)

summary(mod_lm)
```

```{r}
mod_lm %>% 
  autoplot(which = 1:6, label.hjust = 1.2)
```

```{r}
mod_tobit <- tobit(Helpfulness ~ Rating + I(Rating^2) + I(Rating^3)+ log10(Word_Count) + log10(Total_Vote)  + log10(Word_Count)*Rating + Timeline + anticipation  + positive  + negative  + trust,
                   data = movies_data, left = 0, right = 1)

summary(mod_tobit)
```

```{r}
tibble(Residuals = resid(mod_tobit),
       Fitted = fitted(mod_tobit),
       Observed = movies_data$Helpfulness) %>% 
  ggplot(aes(Fitted, Observed))+
  geom_point(alpha = 0.2)+
  geom_abline(slope = 1, color = "red", size = 1.1, linetype = 2)+
  coord_cartesian(xlim = c(0,1), ylim =c(0,1))
```

```{r}
ggplot(data = tibble(Resids = resid(mod_tobit)),
       aes(Resids, ..density..))+
  geom_histogram(bins = 50, color = "white")
```

```{r}
library(GGally)

movies_data %>%
  mutate(Word_Count = log10(Word_Count),
         Total_Vote = log10(Total_Vote),
         n_Sentence = log10(n_Sentence)) %>% 
  select(Helpfulness, Rating, Total_Vote, Word_Count, n_Sentence) %>% 
  ggpairs()
```

```{r}
summary(mod_tobit)

preds <-  predict(mod_tobit, newdata = tibble(Rating = seq(0.5, 5, 0.5),
                                    Word_Count = rep(30,10),
                                    Total_Vote = rep(10,10)))

qplot(x = seq(0.5, 5, 0.5), preds, geom = "point", ylim = c(0,1))
```

```{r}
fitted(mod_tobit) %>% max

movies_data[which.max(fitted(mod_tobit)),]

predict(mod_tobit, newdata = movies_data[which.min(fitted(mod_tobit)),])
```

```{r}
movies_data[which.min(fitted(mod_tobit)),]  %>%
  mutate(Review = str_to_lower(Review, locale = "en")) %>% 
  unnest_tokens(output = word, input = Review, to_lower = FALSE) %>% 
  inner_join(get_sentiments("nrc"))
```