### 2022-05-06 SearchGoods_EDA.Rmd

---
output:
  html_document:
    theme: flatly
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: 4
    df_print: paged
    code_folding: show
---

```{r, include=FALSE}
library(tidyverse)
library(tidytext)
library(scales)
library(AER)
library(ggfortify)
library(lubridate)
library(koRpus.lang.en)

options(scipen = 20)

knitr::opts_chunk$set(warning = FALSE,
                      message = FALSE,
                      fig.align = "center",
                      fig.width = 8)

```

```{r}
## Phone Data

phones_data <- read_csv("all_phone_reviews_15_03_22.csv", show_col_types = FALSE)

phones_data <- phones_data %>% 
  mutate(Product_Type = "Phone")

## Laptop Data

files <- dir()
files <- files[str_detect(dir(), "all_laptop_reviews.+")]
laptop_data <- map_dfr(files, read_csv, show_col_types = FALSE)

laptop_data <- laptop_data %>% 
  filter(!Product_Name %in% c(
    'Samsung - Galaxy Book Pro 360 15.6" AMOLED Touch-Screen Laptop - Intel Evo Platform Core i7 - 16GB Memory - 1TB SSD - Mystic Navy',
    'ASUS - 14.0" Laptop - Intel Celeron N4020 - 4GB Memory - 64GB eMMC - Star Black - Star Black',
    'ASUS - 14" Chromebook - Intel Celeron N3350 - 4GB Memory - 32GB eMMC - Silver',
    'Samsung - Galaxy 13.3" 4K Ultra HD Touch-Screen Chromebook - Intel Core i5 - 8GB Memory - 256GB SSD - Mercury Gray'
  )) %>% 
  mutate(Product_Type = "Laptop")

## TV Data

tvs_data <- read_csv("all_tv_reviews_16_03_22.csv", show_col_types = FALSE)
tvs_data <- tvs_data %>% 
  mutate(Product_Type = "TV")

## Merging All

search_goods <- phones_data %>% 
  bind_rows(tvs_data, laptop_data)

## Filtering Total Votes below 5

search_goods <- search_goods %>% 
  mutate(Total_Vote = Helpful + Unhelpful) %>% 
  filter(Total_Vote > 4) %>% 
  mutate(Helpfulness = Helpful / Total_Vote) %>% 
  mutate(is_Exp = 0)

search_goods <- search_goods %>% 
  mutate(Dupped = duplicated(Review)) %>% 
  filter(Dupped == 0)

## Word Counts

search_goods <- search_goods %>% 
  mutate(Doc_ID = row_number())

words_agg <- search_goods %>%
  mutate(Review = str_to_lower(Review, locale = "en")) %>% 
  unnest_tokens(output = word, input = Review, to_lower = FALSE) %>% 
  #anti_join(stop_words) %>% 
  count(Doc_ID, name = "Word_Count")

search_goods <- search_goods %>% 
  inner_join(words_agg, by = "Doc_ID")

## Emotion Sentiments (EmoLex)

sentiments_agg <- search_goods %>%
  mutate(Review = str_to_lower(Review, locale = "en")) %>% 
  unnest_tokens(output = word, input = Review, to_lower = FALSE) %>% 
  inner_join(get_sentiments("nrc")) %>% 
  count(Doc_ID, sentiment, name = "Sent_Count") %>% 
  pivot_wider(names_from = sentiment, values_from = Sent_Count, values_fill = 0)

search_goods <- search_goods %>% 
  full_join(sentiments_agg, by = "Doc_ID")

variables <- names(sentiments_agg)[-1]

search_goods <- search_goods %>% 
  mutate(across(.cols = variables, .fns = ~ifelse(is.na(.), 0, .) ))

search_goods <- search_goods %>% 
  mutate(across(.cols = all_of(variables), .fns = function(.) ./Word_Count*100))

search_goods <- search_goods %>% 
  mutate(across(.cols = variables, .fns = function(.) log10(. + 1)))

## Dates

search_goods <- search_goods %>% 
  mutate(Date = as_date(mdy_hm(Date, locale = "english")))

first_dates <- search_goods %>% 
  group_by(Product_Name) %>% 
  slice_min(order_by = Date, n = 1, with_ties = F) %>% 
  select(Product_Name, Date) %>% 
  rename(First_Date = Date)

search_goods <- search_goods %>% 
  full_join(first_dates, by = "Product_Name") %>% 
  mutate(Timeline = as.integer(Date - First_Date))

```

```{r}
## Readability Index(Unfortunately wasn't helpful)

# review_vec <-  search_goods %>%
#   pull(Review)
# 
# SMOG_grade <- vector(mode = "double")
# 
# capture.output(
# 
# for (i in seq_len(length(review_vec))) {
#   SMOG <-  readability(tokenize(review_vec[i], lang = "en", format = "obj"),index = "SMOG", quiet = TRUE)
#   
#   SMOG_grade <- append(SMOG_grade, summary(SMOG) %>% pull(grade))
# }
# ,file = nullfile()
# )
# 
# length(SMOG_grade)
# 
# search_goods <- search_goods %>% 
#   mutate(SMOG_grade = as.double(SMOG_grade))
```

# EDA

```{r}
search_goods %>% 
  ggplot(aes(Helpfulness))+
  geom_histogram(color = "white")

search_goods %>% 
  ggplot(aes(Word_Count))+
  geom_histogram(color = "white")+
  scale_x_log10()

search_goods %>% 
  ggplot(aes(Rating))+
  geom_bar(color = "white")

search_goods %>% 
  ggplot(aes(Timeline+1))+
  geom_histogram(color = "white")+
  scale_x_log10()

search_goods %>% 
  ggplot(aes(Total_Vote))+
  geom_histogram(color = "white")+
  scale_x_log10()

```

```{r}
search_goods %>% 
  ggplot(aes(Word_Count, Helpfulness))+
  geom_point()+
  geom_smooth()+
  scale_x_log10()

search_goods %>% 
  ggplot(aes(Word_Count, Helpfulness))+
  geom_point()+
  geom_smooth()+
  facet_wrap(~Rating)+
  scale_x_log10()

search_goods %>% 
  mutate(WC_Binned = cut_number(Word_Count, n = 10)) %>% 
  ggplot(aes(Rating, Helpfulness))+
  geom_point()+
  geom_smooth(method = "loess")+
  facet_wrap(~WC_Binned)

search_goods %>% 
  ggplot(aes(Rating, Helpfulness))+
  geom_point()+
  geom_smooth(method = "loess")

search_goods %>% 
  ggplot(aes(Timeline+1, Helpfulness))+
  geom_point()+
  geom_smooth()+
  scale_x_log10()

search_goods %>% 
  ggplot(aes(Total_Vote, Helpfulness))+
  geom_point()+
  geom_smooth()+
  scale_x_log10()

search_goods %>% 
  select(variables, Word_Count,Rating, Total_Vote, Timeline) %>% 
  GGally::ggcorr(label = T)
```

```{r}
model_lm <- lm(Helpfulness ~ Rating + I(Rating^2) + I(Rating^3)+ log10(Word_Count) + log10(Total_Vote)  + log10(Word_Count)*Rating + Timeline + anticipation  + positive + negative + trust,
               data = search_goods )
summary(model_lm) 
```

```{r}
model_tobit <- tobit(Helpfulness ~ Rating + I(Rating^2) + I(Rating^3)+ log10(Word_Count) + log10(Total_Vote)  + log10(Word_Count)*Rating + Timeline + anticipation   + positive + negative + trust,
                     data = search_goods,
                     left = 0, right = 1)
summary(model_tobit)
```

```{r}
tibble(Residuals = resid(model_tobit),
       Fitted = fitted(model_tobit),
       Observed = search_goods$Helpfulness) %>% 
  ggplot(aes(Fitted, Observed))+
  geom_point(alpha = 0.2)+
  geom_abline(slope = 1, color = "red", size = 1.1, linetype = 2)+
  coord_cartesian(xlim = c(0,1), ylim =c(0,1))
```

```{r}
qplot(resid(model_lm), bins = 50)
qplot(resid(model_tobit), bins = 50)
```

```{r}
toy_data <- tibble(Rating = seq(0.5,5, 0.5),
                   Word_Count = rep(800,10),
                   Total_Vote = rep(100,10))

tibble(
  preds = predict(model_tobit, newdata = toy_data),
  Rating = seq(0.5,5, 0.5)
) %>% 
  ggplot(aes(Rating, preds))+
  geom_line()+
  geom_point()
```